---
title: "Software Engineering Best Practices in R for Data Scientists"
author: "Rodrigo Basa [R Users Group Philippines]"
date: "7/20/2022"
output: 
  ioslides_presentation:
    css: styles.css
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Intro

### Demystify "Software Engineering" 

> Systematic approach to writing good code

## Links

Slides:

https://rug-ph.github.io/code_production_slides


Repository:

https://github.com/RUG-Ph/code_production

## Who is this for? 

- I work alone
- I work in a team of R developers
- I want to share my code with others
- I want to write and publish a package to CRAN
- I need to make my code ready for *production*

## What are we discussing today?

- What? and Why?
- "How about the How?"
- Not arranged in any particular order
- Not an all-encompassing list

## General Don't*s*
- Implement what you don't need
- Try to learn many topics all at once
- Strictly apply everything everywhere

## References For Improving Your Code

- [Advanced R](https://adv-r.hadley.nz)
- [R Packages](https://r-pkgs.org)

## Tools For Your Code

- [library(devtools)](https://devtools.r-lib.org)
- [library(usethis)](https://usethis.r-lib.org/index.html)

## Write Your Code

- Choose one style
  - indent spaces
  - brackets
  - camel case, snake case
- Appropriate object names
  - variable names &rarr; nouns
  - function names &rarr; verbs

## Write Your Code 
### **L1** Style Guides

- [Tidyverse style guide](https://style.tidyverse.org)
- [Google R style guide](https://google.github.io/styleguide/Rguide.html)

## Write Your Code
### **L2** Automate styling

- [library(lintr)](https://lintr.r-lib.org)
- [library(styler)](https://styler.r-lib.org)

Source code `R/01-*.R`

```{r echo=TRUE, eval=FALSE}
lintr::lint("R/01-01_lint_this.R")
styler::style_file("R/01-01_lint_this.R")
```

## Modularize Your Code

- 2000 LOC vs 300 lines of code
- repeated code
- meaningful code
- namespacing &rarr; re-use of object names

Source code `R/02-01_long_code.R`

## Modularize Your Code
### **L1** Functions

- [functions](https://adv-r.hadley.nz/functions.html)

Source code `R/02-*.R`

## Modularize Your Code
### **L2** R6 Classes and Box modules

- [R6 classes](https://r6.r-lib.org/articles/Introduction.html)
- [Box modules](https://klmr.me/box/)

Source code `R/03-*.R`

## Strengthen Your Code

- Stop execution when invalid input is provided.
- GIGO

### External I/O

- [tryCatch()](https://adv-r.hadley.nz/conditions.html)
- [library(checkmate)](https://mllg.github.io/checkmate)

Source code `R/04-*.R`

## Strengthen Your Code

### SQL and other DB queries

- [Safe Queries](https://db.rstudio.com/best-practices/run-queries-safely/)

## Document Your Code

Let others know what your code does. This includes your future self. 

- [library(roxygen2)](https://roxygen2.r-lib.org)
- [vignette](https://r-pkgs.org/vignettes.html)

Source code `R/05-*.R`

## Test Your Code

Make sure your code works as expected.

### **L1** Automate Tests

- [library(testthat)](https://testthat.r-lib.org)
- [library(checkmate)](https://mllg.github.io/checkmate)

Source code `R/06-*.R`

Test code `tests/testthat/`

```{r echo=TRUE, eval=FALSE}
testthat::test_dir("tests/testthat")
```

## Manage Your Code

- data_analysis.R
- data_analysis_1.R
- data_analysis_2.R
- data_analysis_xxxx.R

## Manage Your Code
### Tools

- CLI
- GUI
  - [GitHub Desktop](https://desktop.github.com/)
  - [SourceTree](https://www.sourcetreeapp.com/)
  - [RStudio](https://support.rstudio.com/hc/en-us/articles/200532077-Version-Control-with-Git-and-SVN)

## Manage Your Code
### **L1** Version Control System
- [Git](https://git-scm.com/doc)
- [Github](https://github.com)
- [Bitbucket](https://bitbucket.org)

[Example commit history](https://github.com/RUG-Ph/code_production/commits/main)

## Manage Your Code
### **L2** Advanced VCS
- Branches ([Example](https://github.com/RUG-Ph/code_production/branches))
- Pull requests ([Example](https://github.com/RUG-Ph/code_production/pulls?q=is%3Apr))
- [Git Workflow](https://www.gitkraken.com/learn/git/git-flow)
  - Git Flow
  - Github Flow

## Run Your Code

Manage your run-time environment for reproducibility

- R version
- libraries and their versions

## Run Your Code
### **L1** Renv

- [library(renv)](https://rstudio.github.io/renv/)

Source code `renv.lock`

## Run Your Code
### **L2** Containers

- [Docker](https://www.docker.com/get-started/)

## Test your code Part 2
### **L2** Automate^2^

CICD
- [Github Actions](https://docs.github.com/en/actions) ([Example](https://github.com/RUG-Ph/code_production/actions))

  - [r-lib/actions](https://github.com/r-lib/actions)
  - [Bitbucket Pipelines](https://support.atlassian.com/bitbucket-cloud/docs/get-started-with-bitbucket-pipelines/)
  
Source code `.github/workflows/lint-project.yaml`
